---
title: "data.table tips"
author: "Jordi Puigdellivol"
date: "17 de diciembre de 2015"
output: ioslides_presentation
---

```{r,echo=FALSE}
library(data.table)
```


# Introduction



## what?

Extension of Data.frame
i.e. is.data.frame(DT) == TRUE

 - Think data.table as a set of columns
 - Every column is the same length but different type

```{r}
```


## why?


 - Reduce programming time, fewer function calls, less variable name repetition
 - Fast aggregation of large data (e.g. 100GB in RAM), fast ordered joins, fast add/modify/delete of columns by group using no copies at all, list columns 
 - Fast file reader (fread). 
 - Offers a natural and flexible syntax, for faster development.

```{r}
```


# how?
```{r}
```
## General Form

mtcarsDT[  
<span class="blue">
  mpg > 20
</span>,  
<span class="green">
  .(Avgmean = mean( hp ),
    "minWT(kg)"= min( wt*453.6 ))
</span>,  
<span class="yellow3">
by =.(cyl,under5gear = gear < 5 )
</span>  
]

<div class="columns-2">


<div class="columns-2">
R:  
SQL:  


<span class="blue"> i </span>  
<span class="blue"> where </span>  


</div>  


<div class="columns-2">
<span class="green"> j </span>  
<span class="green"> select </span>  


<span class="yellow3"> by = </span>  
<span class="yellow3"> group by </span>  


</div>


</div>

## Readed

Take <span class="black"> mtcarsDT </span>, subset using <span class="blue"> i </span>, then calculate <span class="green"> j </span> grouped by <span class="yellow3"> by </span>

```{r Readed}
mtcarsDT <- as.data.table(mtcars)
mtcarsDT[
  mpg > 20,
  .(Avgmean = mean( hp ),
    "minWT(kg)"= min( wt*453.6 )),
  by =.(cyl,under5gear = gear < 5 )
  ]
```

## Create

```{r Create}
DT <- data.table( A = 1:6 , B = c("a","b","c") ,
                  C = rnorm(6) , D = TRUE )
DT
```

## Print

```{r Print}
DT2 <- data.table( A = 1:10000000 , B = letters[1:20] ,
                   C = rnorm(10000000) , D = TRUE)
DT2
```

## Subseting

<div class="columns-2">

```{r SubsetingA}
DT[ 3:5 , ] # or DT[ 3:5 ]
```

```{r SubsetingB}
DT[ B == "a" ]
```

```{r SubsetingD}
DT[ B %in% c("a","c") ]
```

```{r SubsetingE}
DT[ between(C,-0.5,0.5) ]
```


</div>

## Selecting columns

```{r selecting}
DT[, .(B, C)]
```

NB: .() is an alias to list() in
data.tables and they mean
the same.

##Computing on columns

<div class="columns-2">
<span>
```{r ComputingA}
DT
```
</span>


<span>
```{r ComputingB}
DT[, .(Total = sum(A),
       Mean = mean(C))]

DT[, .(paste0(B,collapse=""))]
```
  
</span>
</div>

## Recycling
```{r Recycling}
DT[, .(B,C=sum(C))]
```


## Throw anything
```{r Throw_anything}
DT[, plot(A,C)]
```

## Anything
```{r Throw_anything_2}
DT[, { print(A)
       hist(C)
       NULL}]
```








##Performance

```{r perfornamce1,echo=FALSE}
DF <- as.data.frame(DT2)
setkey(DT2,"B")
```

<div class="columns-2">
```{r dimensions}
dim(DF)
```

```{r dimensions2}
dim(DT2)
```

</div>

<div class="columns-2">
```{r perfornamce1A,eval=FALSE,echo=TRUE}
DF[DF$B=="a" ,] 
DT2[.("a"),] )
```

```{r perfornamce1B,eval=TRUE,echo=FALSE}
system.time( DF[DF$B=="a" ,] )
system.time( DT2[.("a"),] )

```


```{r perfornamce2A,eval=FALSE,echo=TRUE}
tapply(DF$C,DF$B,sum)
DT2[,sum(C),by=B]

```

```{r perfornamce2B,eval=TRUE,echo=FALSE}
system.time( tapply(DF$C,DF$B,sum) )
system.time( DT2[,sum(C),by=B] )
```


</div>







#other cool staff


```{r}
```



```{r}
```

## Slide with Plot

```{r, echo=FALSE}
```

# References

***

  - [datacamp](https://www.datacamp.com/courses/data-analysis-the-data-table-way)
  - [user2014](http://user2014.stat.ucla.edu/files/tutorial_Matt.pdf)
  - [Pakage pdf](https://cran.r-project.org/web/packages/data.table/data.table.pdf)
  - [Pakage wiki](https://github.com/Rdatatable/data.table/wiki)
  - [Slides](https://github.com/jpuigde/Rdatatable)

